Estimating the genome-wide contribution of selection to temporal allele frequency change

### 2019-10-08 manuscript/manuscript.tex
updates to ms

\gc{The presence of an unselected control line provides an alternative way to
  partition the effects of linked selection and genetic drift: we can compare
  the total variance in allele frequency change of the control line (which
  excludes the effect of artificial selection on allele frequencies) to the
  total variance in frequency change of the Longshanks selection
  lines.
We can partition the allele frequency change between the two timepoints (20
  generations) for a Longshanks line as $\Delta p_{t,\mathrm{LS1}} =
  \Delta_{_{D1}} p_{t,\mathrm{LS1}} + \Delta_{_{U1}} p_{t} +
  \Delta_{_\mathrm{LS}} p_t$ 
  where these  terms in are the decomposition are the allele frequency change due to
  drift in replicate 1 ($D1$),
  selection unique to the 1$^{st}$ LS replicate ($U1$), and selection response shared between the
  two Longshanks replicates (LS) respectively (and similarly for the Longshanks two line, LS2).
  %partition the allele frequency change between the two timepoints (20
  %generations) for a Longshanks line as $\Delta p_{t,\mathrm{LS1}} =
  % \Delta_{_D} p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} +
  % \Delta_{_\mathrm{LS}} p_t$ (and similarly for the Longshanks two line, LS2),
  % and $\Delta p_{t,\mathrm{Ctrl}} = \Delta_{_D} p_{t,\mathrm{Ctrl}} +
  % \Delta_{_{U}} p_{t,\mathrm{Ctrl}} + \Delta_{_\mathrm{Ctrl}} p_t$ for the
  % control line. Here, $\Delta_{_D} p_{t}$, $\Delta_{_U} p_t$,
  % $\Delta_{_\mathrm{LS}} p_t$ are the allele frequency changes due to drift,
  % selection unique to a replicate, and selection response shared between the
  % two Longshanks replicates, respectively.
  By construction we will assume that each of these terms
  are uncorrelated within replicates, and that only the shared term
  covaries between the replicates. Assuming that we can approximate the
  contribution of genetic drift in the LS
  lines the variance in allele frequency change observed in
  the control, i.e. $\var(\Delta p_{t,\mathrm{Ctrl}})
  = \var(\Delta_{_D1} p_{t})=\var(\Delta_{_D2} p_{t})$, then we can estimate the increase in variance in allele
  frequency change due to selection as $(\var(\Delta p_{t,\mathrm{LS1}}) +
  \var(\Delta p_{t,\mathrm{LS2}}))/2 - \var(\Delta p_{t,\mathrm{Ctrl}})$ and
  the shared effect of selection across selected lines as $\cov(\Delta
  p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$. We estimate at least 32\%
    (95\% CI $[21\%, 48\%]$) of the variance in allele frequency change is
    driven by the effects of selection, of which 14\% (95\% CI $[3\%, 33\%]$)
    is estimated to be unique to a selection line, and 17\% (95\% CI $[9\%,
  23\%]$) is the effect of shared selection between the two Longshanks
selection lines.}

@@

While it challenging to apply temporal methods to natural populations there is
a lot of promise for these approaches
\parencite{Bergland2014-ij,Machado2018-cs}. Efforts to estimate the quantify
the impact of linked selection in have \gc{found over obligately
  sexual origanisms SHOULD WE SAY 0-89} 19\% to 89\% \vb{reduction
in genome-wide diversity} over long time periods is driven by \vb{linked
selection}
\parencite{McVicker2009-ax,Elyashiv2016-vt,Corbett-Detig2015-gt,Coop2016-gx,Comeron2014-nh}
Thus there is reason to be hopeful that we could detect this from temporal
data, which would help to resolve the timescales that linked selection act
over. In our reanalysis of the \textcite{Barghi2019-qy} study, we find evidence
of complex linked selection dynamics, with selection pressures flipping over
time due to either environmental change, the breakup of epistatic combinations
or advantageous haplotypes. Such patterns would be completely obscured in
samples only from contemporary populations, thus we can hope to have a much
richer picture of linked selection as temporal sequencing becomes more common,
\vb{perhaps allowing us to observe the effects of ecological dynamics in
genomic data \parencite{Hairston2005-ga}.}

>>>>

\vince{I think the bulk of this can go to what will become Materials and Methods?}

The presence of an unselected control line provides an alternative way to
partition the effects of linked selection and genetic drift: we can compare the
total variance in allele frequency change of the control line (which excludes
the effect of artificial selection on allele frequencies) to the total variance
in frequency change of the Longshanks selection lines.  We can partition the
allele frequency change between the two timepoints (20 generations) for a
Longshanks line as \vb{$\Delta p_{t,\mathrm{LS1}} = \Delta_{_{D}}
p_{t,\mathrm{LS1}} + \Delta_{_{U}} p_{t,\mathrm{LS1}} + \Delta_{_S}
p_{t,\mathrm{LS}}$} where these  terms in are the decomposition are the allele
frequency change due to drift in \vb{Longshanks replicate 1 ($\Delta_{_D}
p_{t,\mathrm{LS1}}$), selection unique to the LS1 replicate ($\Delta_{_U}
p_{t,\mathrm{LS1}}$), and selection response shared between the two Longshanks
replicates ($\Delta_{_S} p_{t,\mathrm{LS}}$) respectively (and similarly for
the Longshanks two line, LS2)}. By construction we will assume that each of
these terms are uncorrelated within replicates, and that only the shared term
covaries between the replicates. Assuming that we can approximate the
contribution of genetic drift in the LS lines the variance in allele frequency
change observed in the control, i.e. \vb{$\var(\Delta p_{t,\mathrm{Ctrl}}) =
\var(\Delta_{_{D}} p_{t,\mathrm{LS2}})=\var(\Delta_{_{D}}
p_{t,\mathrm{LS2}})$}, then we can estimate the increase in variance in allele
frequency change due to selection as $(\var(\Delta p_{t,\mathrm{LS1}}) +
\var(\Delta p_{t,\mathrm{LS2}}))/2 - \var(\Delta p_{t,\mathrm{Ctrl}})$ and the
shared effect of selection across selected lines as $\cov(\Delta
p_{t,\mathrm{LS1}}, \Delta p_{t,\mathrm{LS2}})$.  We estimate at least 32\%
(95\% CI $[21\%, 48\%]$) of the variance in allele frequency change is driven
by the effects of selection, of which 14\% (95\% CI $[3\%, 33\%]$) is estimated
to be unique to a selection line, and 17\% (95\% CI $[9\%, 23\%]$) is the
effect of shared selection between the two Longshanks selection lines.

@@

While it challenging to apply temporal methods to natural populations there is
a lot of promise for these approaches
\parencite{Bergland2014-ij,Machado2018-cs}. Efforts to estimate the quantify
the impact of linked selection in have found over obligately sexual organisms
\vb{0\% to 89\%} reduction in genome-wide diversity over long time periods is
driven by linked selection
\parencite{McVicker2009-ax,Elyashiv2016-vt,Corbett-Detig2015-gt,Coop2016-gx,Comeron2014-nh}
Thus there is reason to be hopeful that we could detect this from temporal
data, which would help to resolve the timescales that linked selection act
over. In our reanalysis of the \textcite{Barghi2019-qy} study, we find evidence
of complex linked selection dynamics, with selection pressures flipping over
time due to either environmental change, the breakup of epistatic combinations
or advantageous haplotypes. Such patterns would be completely obscured in
samples only from contemporary populations, thus we can hope to have a much
richer picture of linked selection as temporal sequencing becomes more common,
\vb{perhaps allowing us to observe the effects of ecological dynamics in
genomic data \parencite{Hairston2005-ga}.}
